---
title: "Assignment-Two"
output: html_document
date: "2022-10-20"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# U7117486 Assignment Two 

## Load in packages
```{r}
library(pacman)
p_load(bookdown, tidyverse, ggforce, flextable, latex2exp, png, magick, metafor, xtable) 
```

## Reading in the csv file 
```{r}
data <- read_csv("OA_activitydat_20190302_BIOL3207.csv")
data
```

## Omitting N/A data
```{r}
data <- na.omit(data)
data
```

## Drop irrelevant columns
```{r}
data <- subset(data, select = -c(comment, loc))
data
```

## Creating a summary of results 
```{r summaryTab}
summary1<- filter(data, treatment == "control") %>% group_by(species) %>% summarise(mean(activity), sd(activity), n())
colnames(summary1) <- c("Species", "ctrl.mean", "ctrl.sd", "ctrl.n")

summary2<- filter(data, treatment == "CO2") %>% group_by(species) %>% summarise(mean(activity), sd(activity), n())
colnames(summary2) <- c("Species", "oa.mean", "oa.sd", "oa.n")

summary <- merge(summary1, summary2)
summary
```
## Reading in the clark paper meta data
```{r}
clark_meta <- read_csv("clark_paper_data.csv")
clark_meta
```

## Binding the summary results to the study information meta data provided
```{r}
clark_data <- bind_cols(clark_meta, summary)
clark_data
```

## Reading in the larger meta data set 
```{r}
ocean_meta <- read_csv("ocean_meta_data.csv")
ocean_meta
```

## Binding the clark data to the larger meta data set 
```{r}
clark_data[, c(12)] <- sapply(clark_data[, c(12)], as.character)
clark_data[, c(7)] <- sapply(clark_data[, c(7)], as.character)
clark_data[, c(8)] <- sapply(clark_data[, c(8)], as.character)

merged_data <- bind_rows(ocean_meta, clark_data)

colnames(merged_data)[16] <- "Life_Stage"

# Getting rid of negative data
merged_data <- merged_data[which(merged_data$ctrl.n > 0),]
merged_data <- merged_data[which(merged_data$ctrl.mean > 0),]
merged_data <- merged_data[which(merged_data$ctrl.sd > 0),]
merged_data <- merged_data[which(merged_data$oa.n > 0),]
merged_data <- merged_data[which(merged_data$oa.mean > 0),]
merged_data <- merged_data[which(merged_data$oa.sd > 0),]

# Checking that negatives are excluded
max(merged_data$ctrl.mean,na.rm=FALSE)
min(merged_data$ctrl.mean,na.rm=FALSE)
max(merged_data$oa.mean,na.rm=FALSE)
min(merged_data$oa.mean,na.rm=FALSE)
max(merged_data$ctrl.n,na.rm=FALSE)
min(merged_data$ctrl.n,na.rm=FALSE)
max(merged_data$oa.n,na.rm=FALSE)
min(merged_data$oa.n,na.rm=FALSE)
```

## Correctly calculate the log response ratio (lnRR) effect size for every row of the dataframe using metafor’s escalc() function.
```{r}
# Log Response Ratio (lnRR)
merged_data <- metafor::escalc(measure = "ROM", 
                               m1i = ctrl.mean,
                               sd1i = ctrl.sd, 
                               n1i = ctrl.n, 
                               m2i = oa.mean, 
                               sd2i = oa.sd, 
                               n2i = oa.n, 
                               data = merged_data, 
                               var.names = c("lnRR", "vlnRR"))

merged_data$residual <- 1:nrow(merged_data)
```

```{r}
tibble(merged_data)
```
# Correct meta-analytic model fitted to the data that controls for the sampling variance of lnRR. The model should include a random effect of study and observation. Use metafor’s rma.mv() function.

```{r V, echo=TRUE}
## Meta Analytic Model fitted to the data - controls for sampling variance of lnRR.
MLMA <- rma.mv(lnRR ~ 1 + Life_Stage, V = vlnRR,
                        random = list(~1 | Study,
                                      ~1 | residual), 
                        dfs = "contain", 
                        test = "t", 
                        data = merged_data)
summary(MLMA)

## How much variation does this model explain in lnRR? 
r2 <- orchaRd::r2_ml(MLMA)
r2
```

```{r}
## Prediction intervals
predict(MLMA)
```


```{r V, echo=TRUE}
## Meta Analytic Model fitted to the data - controls for sampling variance of lnRR.
MLML <- rma.mv(lnRR ~ 1 + Effect.type , V = vlnRR,
                        random = list(~1 | Study,
                                      ~1 | residual), 
                        dfs = "contain", 
                        test = "t", 
                        data = merged_data)
summary(MLML)

## How much variation does this model explain in lnRR? 
r2 <- orchaRd::r2_ml(MLML)
r2
```

```{r}
# Proportion of total variability when excluding sampling variance
orchaRd::i2_ml(MLMA, data = merged_data)
```

```{r}
## Showing I2 results 

rma(yi = lnRR, 
    vi = vlnRR, 
    method = "DL", 
    data = merged_data)
```

```{r}
On average, there is an approximate 14% decrease in activity of fish for every 1% increase in ocean acidification when compared against the control. We are 95% confident that the true mean falls between -37.13 and 8.65%. There is a significant amount of heterogeneity among the effects (Q = 736,088,769, df = 763, p < 0.0001) with effect sizes expected to be as low as -442.03% and as high as	413.54% 95% of the time (i2 total = 100%). 10% of the model is explained with the r^2 Conditional. The conditional R2 tells us that the full model explains 10% of varaince in effect size, accounting for both fixed and random effects. 

```


## Forest plot showing the mean estimate, 95% confidence interval, and prediction interval with clearly labelled axes, number of samples and studies plotted on figure

```{r}
orchaRd::orchard_plot(MLML, data = merged_data, group = "Study", mod = "Effect.type",
    xlab = "Z-Transformed Correlation Coefficient (Zr)", angle = 45)
```